---
title: "Data Preparation Workflow (Section 2)"
author: "Siyuan"
date: "7/25/2020"
output: 
  html_document:
    fig_width: 10
    fig_height: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, # show code
                      message = FALSE # hide console output
                      )
```

DESeq2 normalization of the gene-level count data

### Step 1: Load the dependencies
```{r}
library(DEE2HsapienData)
library(SummarizedExperiment)
```

### Step 2: Reload the data downloaded from DEE2 in the steps in Section 1
```{r}
tempDirPath <- paste(getwd(), 'temp', sep = '/')
dee2DataPath <- paste(tempDirPath, 'dee2Data', sep='/')
dee2Data <- readRDS(dee2DataPath)
```

### Step 3: Normalize the count data with DESeq2 normalization
```{r}
dESeq2NormalizedData <- Deseq2Normalization(dee2Data)
```

### Step 4: PCA plots
Compare PCA plots of the data before normalization and the data normalized with DESeq2 normalization.

#### PCA plots of the data before normalization

```{r}

pcaPlot(dee2Data$sExpr)
pcaPlot2d <- pcaPlot(dee2Data$sExpr, plot = '2d')
pcaPlot2d[[1]]
pcaPlot2d[[2]]
pcaPlot2d[[3]]
```

#### PCA plots of the data after DESeq2 normalization

```{r}
pcaPlot(dESeq2NormalizedData$normalizedSExpr)
pcaPlot2d <- pcaPlot(dESeq2NormalizedData$normalizedSExpr, 
        plot = '2d')
pcaPlot2d[[1]]
pcaPlot2d[[2]]
pcaPlot2d[[3]]
```

### Step 6: Save the normalized data
```{r}
saveRDS(dESeq2NormalizedData, 
        file=paste(tempDirPath, 'dESeq2NormalizedData', sep='/'))
```