---
title: "Data Preparation Workflow (Section 5)"
author: "Siyuan"
date: "7/25/2020"
output: 
  html_document:
    fig_width: 10
    fig_height: 8
---

TMM Normalization of the gene-level count data summarzied from the transcript-level count data with 'tximport'

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, # show code
                      message = FALSE # hide console output
                      )
```
### Step 1: Load the library
```{r}
## load the library
library(DEE2HsapienData)
library(SummarizedExperiment)
```

### Step 2: Reload the data downloaded from DEE2
```{r}
tempDirPath <- paste(getwd(), 'temp', sep = '/')
dee2DataPath <- paste(tempDirPath, 'dee2Data', sep='/')
dee2Data <- readRDS(dee2DataPath)
```

### Step 3: Normalize the count data
```{r}
tmmNormalizedDataFromTxLevelData <- 
  TMMNormalizationFromTranscriptLevelData(dee2Data)
```

### Step 4: PCA plots
Compare PCA plots of the data before normalization and the normalized data

#### PCA plots of the data before normalization
```{r}
PcaPlot(dee2Data$sExpr)
PcaPlot2d <- PcaPlot(dee2Data$sExpr, plot = '2d')
PcaPlot2d[[1]]
PcaPlot2d[[2]]
PcaPlot2d[[3]]
```

#### PCA plots of the data after TMM normalization
```{r}
PcaPlot(tmmNormalizedDataFromTxLevelData$normalizedSExpr)
PcaPlot2d <- PcaPlot(tmmNormalizedDataFromTxLevelData$normalizedSExpr, 
        plot = '2d')
PcaPlot2d[[1]]
PcaPlot2d[[2]]
PcaPlot2d[[3]]
```
#### Save the normalized data
```{r}
saveRDS(tmmNormalizedDataFromTxLevelData, 
        file=paste(tempDirPath, 'tmmNormalizedDataFromTxLevelData', sep='/'))
```